---
title: "HOLC"
author: "Ruimin"
date: "12/14/2021"
output: html_document
---

```{r}
library(tidyverse)
library(sf)
library(here)
library(fs)
library(stringr)
library(utils)
library(sp)
library(raster)
```


Could work out the % of each HOLC level in current census tracts and then relate that to temperature ...e.g. % of HOLC grade A vs temp, % of HOLC grade B vs temp. 

Basic - just show the temp differences in HOLC neighbourhoods, but then discuss how does this relate to current census tracts

```{r}

census_shape <- st_read(here::here(#"Data", "practice_Q", 
                                  "Census Tracts 2010",
                                  "geo_export_4f0edcf3-ca13-43a5-a50b-879cda3be158.shp"))%>%
  st_transform(., crs=32611) # 

HOLC <- st_read(here::here("CALosAngeles1939", "cartodb-query.shp"))%>%
    st_transform(., crs=32611)
```
Landsat data

```{r eval=F}

listfiles<-dir_info(here::here("Landsat")) %>%
  dplyr::filter(str_detect(path, ".tar")) %>%
  dplyr::select(path)%>%
  dplyr::pull()%>%
  #print out the .gz file
  print()%>%
  as.character()%>%
  utils::untar(exdir=here::here("Landsat"))
```
```{r}
# List your raster files excluding band 8 using the patter argument
Bands_need<-dir_info(here::here("Landsat"))%>%
  dplyr::filter(str_detect(path, "[B456].TIF")) %>%
  #dplyr::filter(str_detect(path, "B10.TIF"))%>%
  #dplyr::filter(str_detect(path, "B11", negate=TRUE))%>%
  dplyr::select(path)%>%
  arrange()%>%
  pull()%>%
  as.character()%>%
  # Load our raster layers into a stack
  raster::stack()

names(Bands_need) <- c('red', 'NIR', 'SWIR1') 

Thermal<-dir_info(here::here("Landsat"))%>%
  dplyr::filter(str_detect(path, "B10.TIF")) %>%
  dplyr::select(path)%>%
  arrange()%>%
  pull()%>%
  as.character()%>%
  # Load our raster layers into a stack
  raster::stack()


thermal_mask <- Thermal %>%
    raster::crop(.,HOLC)

thermal_crop <- thermal_mask %>%
    raster::mask(.,HOLC)


lsatmask1 <- Bands_need %>%
  # now crop our temp data to the extent
  raster::crop(.,HOLC)

# Error in x$.self$finalize() : attempt to apply non-function
# this error is fine to ignore it's from the garbage collector.

raster::plot(thermal_mask)

#takes some time
lsat2 <- lsatmask1%>%
    terra::mask(.,  HOLC)



```
subset the census tracts within HOLC
```{r}
census_sub <- census_shape[HOLC,, op=st_intersects]

library(tmap)


tm_shape(HOLC) +
  tm_polygons(col = "blue")+
  tm_shape(census_sub) +
  tm_polygons(col = "red", alpha = 0.5) 

```



NDVI
```{r}

NDVI <- (lsatmask1$NIR - lsatmask1$red)/ (lsatmask1$NIR + lsatmask1$red)

# you need the sp package loaded for this...

NDVI %>%
  plot(.,col = rev(terrain.colors(10)), main = "Landsat-NDVI")
  

```

MTL - readMTL didn't work due to new projection parameter

```{r}
#library(RStoolbox)

TOA <- (0.0003342 *
  thermal_crop) + 0.1

K1 <- as.numeric(774.8853)

K2 <- as.numeric(1321.0789)
  
Brighttemp <- (K2 / log((K1 / TOA) + 1))

```


```{r}
facveg <- (NDVI-0.2/0.5-0.2)^2

emiss <- (0.004*facveg)+0.986
```

```{r}
Boltzmann <- 1.38*10e-23
Plank <- 6.626*10e-34
c <- 2.998*10e8

p <- Plank*(c/Boltzmann)

lambda <- 1.09e-5

LST <- Brighttemp/(1 +(lambda*Brighttemp/p)*log(emiss))


LST <- LST-273.15


```

NDBI

```{r}

NDBI=((lsatmask1$SWIR1-
         lsatmask1$NIR)/
        (lsatmask1$SWIR1+
        lsatmask1$NIR))

# NDBI %>%
#   raster::writeRaster(., "NDBI", 
#               bylayer=TRUE, 
#               format='GTiff', 
#               overwrite=TRUE)
```

At the miment we have 

* Temp for the extent of HOLC
* Census tracts within the HOLC (intersecting them)

WHat do i want
* Temp within each HOLC
* RElationship between area of HOLC within census tract and temperature so that would mean % of census tract graded A (or B/C/D) and temp

Simple, mean temp per HOLC

```{r}
#LST_per_HOLC <- raster::extract(LST, HOLC, fun=mean, na.rm=TRUE, df=TRUE)

LST_per_census_within_HOLC <- raster::extract(LST, census_sub, fun=mean, na.rm=TRUE, df=TRUE)


LST_per_census_within_HOLC<-LST_per_census_within_HOLC %>%
  mutate(ID=census_sub$ct10)%>%
  rename(temp=layer)


census_sub <- census_sub %>%
  dplyr::mutate(census_area=(st_area(.)))

intersect <- st_intersection(HOLC, census_sub)%>%
    dplyr::mutate(intersect_area=(st_area(.)))

tm_shape(intersect) +
  tm_polygons(col = "blue")+
  tm_shape(HOLC) +
  tm_borders(col="black")

```



```{r}

example_overlap <- intersect %>%
  filter(ct10==125200)

AreaA <- intersect %>%
  group_by(ct10)%>%
  filter(holc_grade=="A")%>%
  summarise(A_area=sum(intersect_area))

AreaB <- intersect %>%
  group_by(ct10)%>%
  filter(holc_grade=="B")%>%
  summarise(B_area=sum(intersect_area))

AreaC <- intersect %>%
  group_by(ct10)%>%
  filter(holc_grade=="C")%>%
  summarise(C_area=sum(intersect_area))

AreaD <- intersect %>%
  group_by(ct10)%>%
  filter(holc_grade=="D")%>%
  summarise(D_area=sum(intersect_area))

```

```{r}

# join_back

census_shapeA <- census_shape %>%
  st_drop_geometry()%>%
  left_join(., 
            AreaA,
            by = c("ct10" = "ct10"))

census_shapeAB <- census_shapeA %>%
  left_join(., 
            AreaB,
            by = c("ct10" = "ct10"))

census_shapeABC <- census_shapeAB %>%
  left_join(., 
            AreaC,
            by = c("ct10" = "ct10"))

census_shapeABCD <- census_shapeABC %>%
  left_join(., 
            AreaD,
            by = c("ct10" = "ct10"))%>%
  as_tibble()


pick<-census_shapeABCD %>%
  dplyr::select(ct10, A_area, B_area, C_area, D_area)

census_shape_joined <- census_shape %>%
  left_join(., 
            pick,
            by = c("ct10" = "ct10"))%>%
  dplyr::mutate(tract_area=st_area(.))


```

% HOLC in tract

```{r}

census_shape_joined_percent <- census_shape_joined %>%
  mutate(percent_area_A=(A_area/tract_area)*100)%>%
  mutate(percent_area_B=(B_area/tract_area)*100)%>%
  mutate(percent_area_C=(C_area/tract_area)*100)%>%
  mutate(percent_area_D=(D_area/tract_area)*100)

```

add in temperature

```{r}
census_shape_joined_percent_temp <- census_shape_joined_percent %>%
  left_join(., 
            LST_per_census_within_HOLC,
            by = c("ct10" = "ID"))

```

```{r}

library(ggplot2)

A <- census_shape_joined_percent_temp %>%
  st_drop_geometry(.)%>%
  #drop_na(.)%>%
  as_tibble(.)%>%
  ggplot(aes(x = as.numeric(percent_area_A) , 
           y = as.numeric(temp)), 
           data=.)+
  geom_point()+
  stat_smooth(method="lm", se=FALSE, size=1)

B <- census_shape_joined_percent_temp %>%
  st_drop_geometry(.)%>%
  #drop_na(.)%>%
  as_tibble(.)%>%
  ggplot(aes(x = as.numeric(percent_area_B) , 
           y = as.numeric(temp)), 
           data=.)+
  geom_point()

C <- census_shape_joined_percent_temp %>%
  st_drop_geometry(.)%>%
  #drop_na(.)%>%
  as_tibble(.)%>%
  ggplot(aes(x = as.numeric(percent_area_C) , 
           y = as.numeric(temp)), 
           data=.)+
  geom_point()


D <- census_shape_joined_percent_temp %>%
  st_drop_geometry(.)%>%
  #drop_na(.)%>%
  as_tibble(.)%>%
  ggplot(aes(x = as.numeric(percent_area_D) , 
           y = as.numeric(temp)), 
           data=.)+
  geom_point()

library(cowplot)

all <- plot_grid(
  A, B, C, D,
  labels = "AUTO", ncol = 2
)

all
```

